Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-15 paper.tex
update

for plain direct-style, the continue function handles more cases.

@@

\begin{lstlisting}
case class VS(vals: Set[Storable], time: Time, store: BStore)
type NCont = Set[VS] => Set[VS]
type Cont  = NCont => (Set[VS] => Set[VS])

def aval(e: Expr, env: Env, store: BStore, time: Time,
          cont: Cont, ncont: NCont): Set[VS] = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]

      def new_ncont(ret_vss: Set[VS])(result: Set[VS]): Set[VS] = {
        val VS(vals, time, store) = ret_vss.head
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aval(e, new_env, new_store, time, (r: NCont) => r, (vss1: Set[VS]) => {
          if (ret_vss.tail.isEmpty) cont(ncont)(result ++ vss1)
          else new_ncont(ret_vss.tail)(result ++ vss1)
        })
      }

      def nd(closures: List[Clos], vss_acc: Set[VS]): Set[VS] = {
        val Clos(Lam(v, body), c_env)::rest = closures
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, aeval(ae, env, store))
        aval(body, new_env, new_store, new_time, (r: NCont) => r, (vss: Set[VS]) => {
          if (rest.isEmpty) new_ncont(vss_acc ++ vss)(Set())
          else nd(rest, vss_acc ++ vss)
        })
      }

      nd(closures, Set())

    case ae if isAtomic(ae) =>
      val ds = aeval(ae, env, store)
      cont(ncont)(Set(VS(ds, new_time, store)))
  }
}

@@

def analyze(e: Expr) = aval(e, mtEnv, mtStore, mtTime, 
                            (r: NCont) => (vss: Set[VS]) =>  r(vss), 
                            (vss: Set[VS]) => vss)
\end{lstlisting}

@@

\begin{lstlisting}
\end{lstlisting}

@@

\begin{lstlisting}
def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): (Set[VS], Cache) = {
  val config = Config(e, env, time)
  if (cache.outContains(config)) return (cache.outGet(config), cache)

  e match {
    case Let(x, App(f, ae), e) =>
      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
      val app_vss_cache = for (Clos(Lam(v, body), c_env) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, aeval(ae, env, store))
        aval(body, new_env, new_store, new_time, new_cache) //TODO: Do we have to use the latest cache?
      }

      val app_vss = app_vss_cache.map(_._1).foldLeft(Set[VS]())(_ ++ _)
      val app_cache = app_vss_cache.map(_._2).foldLeft(mtCache)(_.outJoin(_))

      val e_vss = for (VS(vals, time, store) <- app_vss.toList) yield {
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aval(e, new_env, new_store, time, app_cache) //TODO: Do we have to use the latest cache?
      }
      val result = e_vss.map(_._1).foldLeft(Set[VS]())(_ ++ _)
      (result, e_vss.map(_._2).foldLeft(mtCache)(_.outJoin(_)).outUpdate(config, result))

    case ae if isAtomic(ae) =>
      val ds = aeval(ae, env, store)
      (Set(VS(ds, new_time, store)), new_cache.outUpdate(config, VS(ds, new_time, store)))
  }
}

def analyze(e: Expr) = {
  def iter(cache: Cache): (Set[VS], Cache) = {
    val (vss, new_cache) = aval(e, mtEnv, mtStore, mtTime, cache)
    if (new_cache.out == cache.out) (vss, new_cache)
    else iter (Cache(new_cache.out, new_cache.out))
  }
  iter(mtCache)
}
\end{lstlisting}

>>>>

for plain direct-style, the continue function handles more cases.
\todo{a defunctionalized object for non-determinism}

@@

\begin{lstlisting}
def nd[T,S](ts: Set[T], acc: S, f: (T, S, S=>S) => S, g: S => S): S = {
  if (ts.isEmpty) g(acc)
  else f(ts.head, acc, (vss: S) => nd(ts.tail, vss, f, g))
}

type Ans = Set[VS]
type Cont = Ans => Ans

def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont): Ans = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
      nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
        val Clos(Lam(v, body), c_env) = clos
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, atomicEval(ae, env, store)) 
        aval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
          nd[VS, Ans](bodyvss, Set[VS](), { case (vs, acc_vss, bdnd) => 
            val VS(vals, time, store) = vs
            val new_env = env + (x -> baddr)
            val new_store = store.update(baddr, vals)
            aval(e, new_env, new_store, time, (evss: Ans) => bdnd(acc_vss ++ evss))
          },
          (evss: Ans) => closnd(evss ++ acc))
        })
      }, 
      cont)
    case ae if isAtomic(ae) => 
      val ds = atomicEval(ae, env, store)
      cont(Set(VS(ds, new_time, store))) 
  } 
}

@@

def analyze(e: Expr) =
  aval(e, mtEnv, mtStore, mtTime, (ans => ans))
\end{lstlisting}

@@

\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time)
case class Cache(in: Store[Config, VS], out: Store[Config, VS]) {
  def inGet(config: Config): Set[VS] = in.getOrElse(config, Set())
  def outGet(config: Config): Set[VS] = out.getOrElse(config, Set())
  def outContains(config: Config): Boolean = out.contains(config)
  def outUpdate(config: Config, vss: Set[VS]): Cache = Cache(in, out.update(config, vss))
}
\end{lstlisting}

adding cache to eval

The definition of Ans is changed to a case class that contains a set of @VS@ and
a cache.
\begin{lstlisting}
case class Ans(vss: Set[VS], cache: Cache)
\end{lstlisting}

\begin{lstlisting}
def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) return cont(Ans(cache.outGet(config), cache))

  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>
        val Clos(Lam(v, body), c_env) = clos
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, atomicEval(ae, env, store))
        aval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
          nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
            val VS(vals, time, store) = vs
            val baddr = allocBind(x, time)
            val new_env = env + (x -> baddr)
            val new_store = store.update(baddr, vals)
            aval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => bdnd(Ans(acc_vss ++ evss, ecache)) })
          },
          { case Ans(evss, cache) => closnd(Ans(evss ++ acc, cache)) })
        })
      },
      { case Ans(vss, cache) => cont(Ans(vss, cache.outUpdate(config, vss))) })

    case ae if isAtomic(ae) =>
      val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
      cont(Ans(vs, new_cache.outUpdate(config, vs)))
  }
}
def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, new_cache) = aval(e, env, store, mtTime, cache, { 
      val initConfig = Config(e, env, store, mtTime)
      case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss)) 
    })
    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
    else { iter(Cache(new_cache.out, new_cache.out)) }
  }    
  iter(mtCache)
}
\end{lstlisting}

@@

\begin{lstlisting}
def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) {
    return Ans(cache.outGet(config), cache)
  }

  var new_cache = cache.outUpdate(config, cache.inGet(config))
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
      val letvs = for (Clos(Lam(v, body), c_env) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, aeval(ae, env, store))
        val Ans(bdv, bdcache) = aval(body, new_env, new_store, new_time, new_cache)
        new_cache = bdcache
        val evs = for (VS(vals, time, store) <- bdv) yield {
          val baddr = allocBind(x, time)
          val new_env = env + (x -> baddr)
          val new_store = store.update(baddr, vals)
          val Ans(ev, ecache) = aval(e, new_env, new_store, time, new_cache)
          new_cache = ecache
          ev
        }
        evs.foldLeft(Set[VS]())(_ ++ _)
      }
      val letv = letvs.foldLeft(Set[VS]())(_ ++ _)
      Ans(letv, new_cache.outUpdate(config, letv))

    case ae if isAtomic(ae) =>
      val vs = Set(VS(aeval(ae, env, store), new_time, store))
      Ans(vs, new_cache.outUpdate(config, vs))
  }
}
\end{lstlisting}
